Key Derivation From Noisy Sources With More Errors Than Entropy

### 2014-12-01 constantRate/ConstCompFuzzExt-reorg.tex
minor typo fixes and grant additions

\begin{construction}
\label{cons:info theoretic}
Let $\mathcal{Z}$ be an alphabet and let $W=W_1,..., W_\gamma$ be a distribution over $\mathcal{Z}^\gamma$.  We describe $\gen, \rep$ as follows:
\begin{center}
\begin{tabular}{c|c}
\begin{minipage}{3in}
\textbf{\gen}
\begin{enumerate}
\item \underline{Input}: $w = w_1,..., w_\gamma$
\item For $j=1,..., \gamma$:
\begin{enumerate}[(i)]
\item Sample $seed_i\leftarrow \zo^d$.
\item Set $v_i = \cond(w_i, seed_i)$.
\end{enumerate}
\item Set $(r, p') \leftarrow \gen'(v_1,..., v_\gamma)$.
\item Set $p = (p', seed_1,..., seed_\gamma)$.
\item Output $(r, p)$.
\end{enumerate}
 \end{minipage} &
\begin{minipage}{3in}
\textbf{\rep}
\begin{enumerate}
\item \underline{Input}: $(w', p = (p', seed_1,..., seed_\gamma))$
\item For $j=1,..., \gamma$:
\begin{enumerate}[(i)]
\item Set $v_i = \cond(w_i, seed_i)$.
\end{enumerate}
\item Output $r = \rep'(v', ss)$.
\end{enumerate}
\vspace{0.7in}
\end{minipage}
\end{tabular}
\end{center}
\end{construction}

@@

\begin{definition}
\label{def:partial source}
A distribution $W = W_1,..., W_\gamma$ is an $(\alpha, \beta)$-partial block source if there exists a set of indices $J$ where $|J| \geq \gamma - \beta$ such that the following holds:
\[
\forall j\in J, \forall w_1,..., w_\gamma \in W_1,..., W_\gamma, \Hoo(W_j | W_1 = w_1,..., W_{j-1}=w_{j-1}) \geq \alpha.
\]
\end{definition}
\defref{def:partial source} is a weakening of block sources~(introduced by Chor and Goldreich~\cite{DBLP:journals/siamcomp/ChorG88}), as only some blocks are required to have entropy conditioned on the past.  The choice of conditioning on the past is arbitrary: a more general sufficient condition is that there exists some ordering of indices where most items have entropy conditioned on all previous items in this ordering~(for example, a ``partial'' reverse block source~\cite{vadhan2003constructing}).  This construction is secure and it supports distributions with more errors than entropy~(proof is in~\apref{sec:info theory sec}).

\begin{lemma}
\label{lem:info theory sec}
%Let $n$ be a security parameter and let $\gamma = \omega(\log n)$.
Let $\mathcal{W}$ be the family of $(\alpha = \Omega(1), \beta\leq \gamma(1-\Theta(1)))$-partial block sources over $\mathcal{Z}^\gamma$ and let $\cond: \mathcal{Z} \times \zo^d\rightarrow \mathcal{Y}$ be a $(\alpha, \tilde{\alpha}, \epsilon_{cond})$-randomness conductor.  Let $(\gen', \rep')$ be $(\mathcal{Y}^\gamma, \tilde{\alpha}(\gamma - \beta), \kappa, t, \epsilon_{fext})$-fuzzy extractor with error $\delta$.\footnote{We actually need $(\gen', \rep')$ to be an average case fuzzy extractor~(see \cite[Definition 4]{DBLP:journals/siamcomp/DodisORS08} and the accompanying discussion).  Most known constructions of fuzzy extractors are average-case fuzzy extractors.  For simplicity we refer to $\gen', \rep'$ as simply a fuzzy extractor.}  Then $(\gen, \rep)$ is a $(\mathcal{Z}^\gamma, \mathcal{W}, \kappa, t, \gamma\epsilon_{cond}+\epsilon_{fext})$-fuzzy extractor with error $\delta$.

@@

\begin{center}
\begin{tabular}{c|c}
\begin{minipage}{3in}
\textbf{\gen}
\begin{enumerate}
\item \underline{Input}: $w = w_1,..., w_\gamma$
\item Sample $r \overset{\$}\leftarrow \zo^\kappa$.
\item For $i=1,..., \ell$:
\begin{enumerate}[(i)]
\item Select $\lambda_i\overset{\$}\leftarrow \zo^{r_{sam}}$.
\item Set $j_{i, 1},..., j_{i, \eta}\leftarrow \sample_{\gamma,\eta}( \lambda_i)$
\item Set $v_i = w_{j_{i,1}},..., w_{j_{i, \eta}}$.
\item Set $\rho_i = \mathcal{O}(I_{v_i, r})$.
\item Set $p_i = \rho_i, \lambda_i$.
\end{enumerate}
\item Output $(r, p)$, where $p=p_1\dots p_\ell$.
\end{enumerate}
 \end{minipage} &
\begin{minipage}{3in}
\textbf{\rep}
\begin{enumerate}
\item \underline{Input}: $(w', p)$
\item For $i=1,..., \ell$:
\begin{enumerate}[(i)]
\item Parse $p_i$ as $\rho_i, \lambda_i$.
\item Set $j_{i, 1},..., j_{i, \eta}\leftarrow \sample_{\gamma, \eta}(\lambda_i)$.
\item Set $v_i' = w_{j_{i, 1}},..., w_{j_{i, \eta}}$.
\item Set $\rho_i(v_i') = r_i$.  If $r_i\neq \perp$ output $r_i$.
\end{enumerate}
\item Output $\perp$.
\end{enumerate}
\vspace{0.37in}
\end{minipage}
\end{tabular}
\end{center}
\end{construction}

@@

\noindent
We can then argue that all $V_i$ simultaneously have individual entropy with good probability~(by union bound):
\begin{corollary}
\label{cor:samp sec}
 Let $\epsilon_{sam}, \alpha'$ be as in \lemref{lem:sampling works}, and all the other variables be as in \thref{thm:sampling}.  Then $\Pr_{\lambda\leftarrow \Lambda}[\forall i, \Hoo(V_i | \Lambda = \lambda)  \ge \alpha'] \leq 1-\ell\epsilon_{sam}$.
%[V_i(V, \Lambda)$ is $(\ell\epsilon_{sam})$-close to a distribution $(V', U_{\ell\times r_{sam}})$ where for $u\in U_{\ell\times r_{sam}}$ for all $i$, $\Hoo(V_i' | U_{\ell\times r_{sam}} =u)\geq \alpha'$.
\end{corollary}
%\begin{proof}
%Union bound over the probability in \lemref{lem:sampling works}.
%%Hybrid argument over the statistical distance in \lemref{lem:sampling works}.
%\end{proof}

@@

Ran Canetti is supported by the NSF MACS project, an NSF Algorithmic foundations grant 1218461, the Check Point Institute for Information Security, and  ISF grant 1523/14.
Omer Paneth is additionally supported by the Simons award for graduate students in theoretical computer science.
The work of Benjamin Fuller is sponsored in part by US NSF grants 1012910 and 1012798 and  the United States Air Force under Air Force Contract FA8721-05-C-0002. Opinions, interpretations, conclusions and recommendations are those of the authors and are not necessarily endorsed by the United States Government. 
Leonid Reyzin is supported in part by US NSF grants 0831281, 1012910, 1012798, and 1422965.
}
\bibliographystyle{alpha}
\bibliography{crypto}

@@

We now argue that there exists a distribution $Y$ where $\Hav(Y | seed)\ge \tilde{\alpha}(\gamma - \beta)$ and $(V, seed_1,..., seed_\gamma)\approx (Y, seed_1,.., seed_\gamma)$.  First note since $W$ is $(\alpha, \beta)$-partial block distribution that
there exists a set of indices $J$ where $|J| \geq \gamma - \beta$ such that the following holds:
\[
\forall j\in J, \forall w_1,..., w_\gamma \in W_1,..., W_\gamma, \Hoo(W_j | W_1 = w_1,..., W_{j-1}=w_{j-1}) \geq \alpha.
\]
Then consider the first element of $j_1\in J$, $\forall w_1,..., w_{j_1-1}\in W_1,..., W_{j_1-1}$,
\[\Hoo(W_{j_1} | W_1 = w_1,..., W_{j_1-1} = w_{j_1-1})\ge \alpha.\]

%\[(\ext (W_1, seed_1), seed_1) \approx_{\epsilon} (U_{\mathcal{Y}}, seed_1)\]
%Then since $W$ is a block-source,
%\[ \forall w_1\in W_1, \Hoo(W_2 | w_1)\ge k.\]
\noindent
Thus, there exists a distribution $Y_{j_1}$ with $\Hav(Y_{j_1} | seed_{j_1}) \ge \tilde{\alpha}$ such that
\[(\cond (W_{j_1}, seed_{j_1}), seed_{j_1}, W_1,..., W_{j_1-}) \approx_{\epsilon_{cond}} (Y_{j_1}, seed_{j_1}, W_1,..., W_{j_1})\]
and since $(seed_1,..., seed_{j_1})$ are independent of these values
\[(\cond (W_{j_1},seed_{j_1}), W_{j_1-1},..., W_1, seed_{j_1}, ..., seed_{1}) \approx_{\epsilon_{cond}} (Y_{j_1}, W_{j_1-1},..., W_1, seed_{j_1}, , ...,  seed_{1})\]
consider the random variable $Z_{j_1} =( Y_{j_1}, \cond(W_{j_1-1},seed_{j_1-1}),..., \cond(W_{1}, seed_{1}))$ and note that \[\Hav(Z_{j_1} | seed_1,...,seed_{j_1})\ge \alpha'.\]
Applying a deterministic function does not increase statistical distance and thus,
\begin{align*}
(\cond (W_{j_1}, seed_{j_1}), \cond(W_{j_1-1}, seed_{j_1-1}),..., \cond(W_1, seed_1), seed_{j_1},..., seed_{1}) \\\approx_{\gamma \epsilon_{cond}} (Z_{j_1}, seed_{j_1},..., seed_1)
\end{align*}

>>>>

\begin{construction}
\label{cons:info theoretic}
Let $\mathcal{Z}$ be an alphabet and let $W=W_1,..., W_\gamma$ be a distribution over $\mathcal{Z}^\gamma$.  We describe $\gen, \rep$ as follows:
\begin{center}
\begin{tabular}{c|c}
\begin{minipage}{3in}
\textbf{\gen}
\begin{enumerate}
\item \underline{Input}: $w = w_1,..., w_\gamma$
\item For $j=1,..., \gamma$:
\begin{enumerate}[(i)]
\item Sample $seed_i\leftarrow \zo^d$.
\item Set $v_i = \cond(w_i, seed_i)$.
\end{enumerate}
\item Set $(r, p') \leftarrow \gen'(v_1,..., v_\gamma)$.
\item Set $p = (p', seed_1,..., seed_\gamma)$.
\item Output $(r, p)$.
\end{enumerate}
 \end{minipage} &
\begin{minipage}{3in}
\textbf{\rep}
\begin{enumerate}
\item \underline{Input}: $(w', p = (p', seed_1,..., seed_\gamma))$
\item For $j=1,..., \gamma$:
\begin{enumerate}[(i)]
\item Set $v_i' = \cond(w_i', seed_i)$.
\end{enumerate}
\item Output $r = \rep'(v', p')$.
\end{enumerate}
\vspace{0.7in}
\end{minipage}
\end{tabular}
\end{center}
\end{construction}

@@

\begin{definition}
\label{def:partial source}
A distribution $W = W_1,..., W_\gamma$ is an $(\alpha, \beta)$-partial block source if there exists a set of indices $J$ where $|J| \geq \gamma - \beta$ such that the following holds:
\[
\forall j\in J, \forall w_1,..., w_{j-1} \in W_1,..., W_{j-1}, \Hoo(W_j | W_1 = w_1,..., W_{j-1}=w_{j-1}) \geq \alpha.
\]
\end{definition}
\defref{def:partial source} is a weakening of block sources~(introduced by Chor and Goldreich~\cite{DBLP:journals/siamcomp/ChorG88}), as only some blocks are required to have entropy conditioned on the past.  The choice of conditioning on the past is arbitrary: a more general sufficient condition is that there exists some ordering of indices where most items have entropy conditioned on all previous items in this ordering~(for example, a ``partial'' reverse block source~\cite{vadhan2003constructing}).  This construction is secure and it supports distributions with more errors than entropy~(proof is in~\apref{sec:info theory sec}).

\begin{lemma}
\label{lem:info theory sec}
%Let $n$ be a security parameter and let $\gamma = \omega(\log n)$.
Let $\mathcal{W}$ be the family of $(\alpha = \Omega(1), \beta\leq \gamma(1-\Theta(1)))$-partial block sources over $\mathcal{Z}^\gamma$ and let $\cond: \mathcal{Z} \times \zo^d\rightarrow \mathcal{Y}$ be a $(\alpha, \tilde{\alpha}, \epsilon_{cond})$-randomness conductor.  Define $\mathcal{V}$ as the family of all distributions with min-entropy at least $\tilde{\alpha}(\gamma-\beta)$ and let $(\gen', \rep')$ be $(\mathcal{Y}^\gamma, \mathcal{V}, \kappa, t, \epsilon_{fext})$-fuzzy extractor with error $\delta$.\footnote{We actually need $(\gen', \rep')$ to be an average case fuzzy extractor~(see \cite[Definition 4]{DBLP:journals/siamcomp/DodisORS08} and the accompanying discussion).  Most known constructions of fuzzy extractors are average-case fuzzy extractors.  For simplicity we refer to $\gen', \rep'$ as simply a fuzzy extractor.}  Then $(\gen, \rep)$ is a $(\mathcal{Z}^\gamma, \mathcal{W}, \kappa, t, \gamma\epsilon_{cond}+\epsilon_{fext})$-fuzzy extractor with error $\delta$.

@@

\begin{center}
\begin{tabular}{c|c}
\begin{minipage}{3in}
\textbf{\gen}
\begin{enumerate}
\item \underline{Input}: $w = w_1,..., w_\gamma$
\item Sample $r \overset{\$}\leftarrow \zo^\kappa$.
\item For $i=1,..., \ell$:
\begin{enumerate}[(i)]
\item Select $\lambda_i\overset{\$}\leftarrow \zo^{r_{sam}}$.
\item Set $j_{i, 1},..., j_{i, \eta}\leftarrow \sample_{\gamma,\eta}( \lambda_i)$
\item Set $v_i = w_{j_{i,1}},..., w_{j_{i, \eta}}$.
\item Set $\rho_i = \mathcal{O}(I_{v_i, r})$.
\item Set $p_i = \rho_i, \lambda_i$.
\end{enumerate}
\item Output $(r, p)$, where $p=p_1\dots p_\ell$.
\end{enumerate}
 \end{minipage} &
\begin{minipage}{3in}
\textbf{\rep}
\begin{enumerate}
\item \underline{Input}: $(w'=w_1',..., w_\gamma', p)$
\item For $i=1,..., \ell$:
\begin{enumerate}[(i)]
\item Parse $p_i$ as $\rho_i, \lambda_i$.
\item Set $j_{i, 1},..., j_{i, \eta}\leftarrow \sample_{\gamma, \eta}(\lambda_i)$.
\item Set $v_i' = w_{j_{i, 1}}',..., w_{j_{i, \eta}}'$.
\item Set $\rho_i(v_i') = r_i$.  If $r_i\neq \perp$ output $r_i$.
\end{enumerate}
\item Output $\perp$.
\end{enumerate}
\vspace{0.37in}
\end{minipage}
\end{tabular}
\end{center}
\end{construction}

@@

\noindent
We can then argue that all $V_i$ simultaneously have individual entropy with good probability~(by union bound):
\begin{corollary}
\label{cor:samp sec}
 Let $\epsilon_{sam}, \alpha'$ be as in \lemref{lem:sampling works}, and all the other variables be as in \thref{thm:sampling}.  Then $\Pr_{\lambda\leftarrow \Lambda}[\forall i, \Hoo(V_i | \Lambda = \lambda)  \ge \alpha'] \geq 1-\ell\epsilon_{sam}$.
%[V_i(V, \Lambda)$ is $(\ell\epsilon_{sam})$-close to a distribution $(V', U_{\ell\times r_{sam}})$ where for $u\in U_{\ell\times r_{sam}}$ for all $i$, $\Hoo(V_i' | U_{\ell\times r_{sam}} =u)\geq \alpha'$.
\end{corollary}
%\begin{proof}
%Union bound over the probability in \lemref{lem:sampling works}.
%%Hybrid argument over the statistical distance in \lemref{lem:sampling works}.
%\end{proof}

@@

Ran Canetti is supported by the NSF MACS project, an NSF Algorithmic foundations grant 1218461, the Check Point Institute for Information Security, and  ISF grant 1523/14.
Omer Paneth is additionally supported by the Simons award for graduate students in theoretical computer science.
The work of Benjamin Fuller is sponsored in part by US NSF grants 1012910 and 1012798 and  the United States Air Force under Air Force Contract FA8721-05-C-0002. Opinions, interpretations, conclusions and recommendations are those of the authors and are not necessarily endorsed by the United States Government. 
Leonid Reyzin is supported in part by US NSF grants 0831281, 1012910, 1012798, and 1422965.  Adam Smith is supported in part by NSF awards 0747294 and 0941553. 
}
\bibliographystyle{alpha}
\bibliography{crypto}

@@

We now argue that there exists a distribution $Y$ where $\Hav(Y | seed)\ge \tilde{\alpha}(\gamma - \beta)$ and $(V, seed_1,..., seed_\gamma)\approx (Y, seed_1,.., seed_\gamma)$.  First note since $W$ is $(\alpha, \beta)$-partial block distribution that
there exists a set of indices $J$ where $|J| \geq \gamma - \beta$ such that the following holds:
\[
\forall j\in J, \forall w_1,..., w_{j-1} \in W_1,..., W_{j-1}, \Hoo(W_j | W_1 = w_1,..., W_{j-1}=w_{j-1}) \geq \alpha.
\]
Then consider the first element of $j_1\in J$, $\forall w_1,..., w_{j_1-1}\in W_1,..., W_{j_1-1}$,
\[\Hoo(W_{j_1} | W_1 = w_1,..., W_{j_1-1} = w_{j_1-1})\ge \alpha.\]

%\[(\ext (W_1, seed_1), seed_1) \approx_{\epsilon} (U_{\mathcal{Y}}, seed_1)\]
%Then since $W$ is a block-source,
%\[ \forall w_1\in W_1, \Hoo(W_2 | w_1)\ge k.\]
\noindent
Thus, there exists a distribution $Y_{j_1}$ with $\Hav(Y_{j_1} | seed_{j_1}) \ge \tilde{\alpha}$ such that
\[(\cond (W_{j_1}, seed_{j_1}), seed_{j_1}, W_1,..., W_{j_1-1}) \approx_{\epsilon_{cond}} (Y_{j_1}, seed_{j_1}, W_1,..., W_{j_1-1})\]
and since $(seed_1,..., seed_{j_1})$ are independent of these values
\[(\cond (W_{j_1},seed_{j_1}), W_{j_1-1},..., W_1, seed_{j_1}, ..., seed_{1}) \approx_{\epsilon_{cond}} (Y_{j_1}, W_{j_1-1},..., W_1, seed_{j_1}, , ...,  seed_{1})\]
consider the random variable $Z_{j_1} =( Y_{j_1}, \cond(W_{j_1-1},seed_{j_1-1}),..., \cond(W_{1}, seed_{1}))$ and note that \[\Hav(Z_{j_1} | seed_1,...,seed_{j_1})\ge \alpha'.\]
Applying a deterministic function does not increase statistical distance and thus,
\begin{align*}
(\cond (W_{j_1}, seed_{j_1}), \cond(W_{j_1-1}, seed_{j_1-1}),..., \cond(W_1, seed_1), seed_{j_1},..., seed_{1}) \\\approx_{\gamma \epsilon_{cond}} (Z_{j_1}, seed_{j_1},..., seed_1)
\end{align*}
